Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-16 paper/manuscript.tex
working backwards with fixes

In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of alchemical calculations~(\ref{sec:step3}). 

@@

@@

Based on the input ligand series a perturbation map or network can be planned. Recent heuristics have shown the more connected the perturbation network the better, however, there is a way to optimise network structure while minimising the number of perturbations that need to be computed minimising the resulting computational cost~\cite{yang2019optimal}. Sometimes the introduction of intermediates that are not part of the original congeneric series are essential to avoid ring breaking, or dealing with perturbations that would otherwise result in large numbers of atoms to be grown or disappearing. Some commercial tools have underlying good heuristic but may fail with complicated input, needing user validation in particular when dealing with chiral compounds. 

@@

Methods in experimental design have been applied to the construction of the perturbation maps.  Yang et al.~\cite{yang2019optimal} optimized the perturbation map by selecting a fixed number of calculations from the pairwise perturbations so that the resulting set of calculations minimize the total variance.  Xu~\cite{xu2019diffnet} optimized the perturbation map by allocating different amounts of simulation time to different pairwise perturbations so as to minimize the total variance, given the total simulation time of all the perturbation calculations.  Both approaches lead to substantial reduction in the statistical error of the estimated free energies.  

@@

\subsubsection{The importance of the alchemical pathway\label{sec:important_path}}

@@

Free energy information should generally be saved more frequently than coordinate data, approximately at the rate that uncorrelated samples are produced.  
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected frequently. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. One may safely assume that the correlation time is greater than 100-200 fs even for relatively simple systems such as small molecules in solvent, so saving no more frequently than every 50-100 steps is recommended. 
How decorrelation impacts calculations, and how to compute it is discussed in other sections.  [section ref]

\section{Step 3 -- Data analysis}
\label{sec:step3}

Once data has been collected from alchemical intermediates, it must be analyzed to produce an estimate of the free energy change (and its associated statistical uncertainty) for each leg of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to provide information on poor convergence.

\subsection{Detecting the boundary between equilibrated and production regions}
\label{sec:automatic-equilibration-detection}

Much of the infrastructure for analyzing alchemical free energy calculations relies on the concept of asymptotically unbiased estimators, which produce unbiased estimates of the free energy when fed very long simulations~\cite{shirts2005comparison}.
In reality, free energy calculations are often initiated from highly atypical initial conditions (such as a protein-ligand geometry obtained from docking and subjected to a heuristic solvent placement scheme), and simulations are of a finite length dictated by available computational resources and computing demands.
As a result, these estimators can produce significantly biased estimates if fed the entirety of simulation data generated without further processing~\cite{chodera2016simple}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/fig11_equib_detection/Figure.pdf}
    \caption{{\bf Automatic partitioning into equilibration and production regions.}
    \textbf{A} The average (black line) standard deviation (shaded region) of the reduced potential $u^*$ over many independent replicate simulations started from the same initial conditions show a significant initial transient before relaxing to the true average potential energy\textbf{B} A cumulative average (red) of the entire simulation data demonstrates simulation bias not seen when initial simulation data is omitted (blue). Using an automated approach to detect equilibration of the boundary $t_0$ using statistical inefficiency $g$ (\textbf{C})  for an effective simulation interval (\textbf{D}) \textbf{E} The optimal equilibration boundary $t_0$ is selected to maximize the number of uncorrelated samples.
    \emph{Figure adapted from~\cite{chodera2016simple}.}
    }
    \label{fig:automatic-equilibration-detection}
\end{figure}

@@

\paragraph{Cautions in automating equilibration detection}
For simulations that are simply not long enough to contain a large number of samples from true equilibrium either because they are very short or contain slow processes, this procedure cannot completely remove the bias.
In such cases, this approach simply selects the final portion of the the simulation, which may be contained in a single substate of conformational space, which may itself lead to biased estimates. 
This situation can be detected if the equilibration boundary $t_0$ is a significant fraction of the total simulation length $T$, with a good rule of thumb being that $T \gtrsim 20 t_0$.
If this is not possible, advanced analysis techniques that assume only local equilibrium (rather than global equilibrium) such as the TRAM estimators~\cite{mey2014xtram,wu2016multiensemble,nuske2017markov} may be more appropriate.

\subsection{Decorrelating samples for analysis}
\label{sec:decorrelating-samples}

\paragraph{Computing the statistical inefficiency}
Most estimators require an uncorrelated set of samples from the equilibrium distribution to produce (relatively) unbiased estimates of the free energy difference and its statistical uncertainty.
To do this, the production region of the simulation is generally \emph{subsampled} with an interval approximately equal to or greater than the \emph{statistical inefficiency} $g \ge 1$ to produce a set of uncorrelated samples that can be fed to the estimator machinery~\cite{chodera2016simple},
\begin{eqnarray}
g &\equiv& 1 + 2 \tau_\mathrm{eq} \label{eq:statistical-inefficiency-definition}
\end{eqnarray}
where $\tau_\mathrm{eq}$ is the integrated autocorrelation time, formally defined as
\begin{eqnarray}
\tau_\mathrm{eq} &\equiv& \sum_{t=1}^{T-1} \left(1 - \frac{t}{T}\right) C_t \label{eq:integrated-autocorrelation-time-definition} , 
\end{eqnarray}
with the discrete-time normalized fluctuation autocorrelation function $C_t$ defined as
\begin{eqnarray}
C_t &\equiv& \frac{\expect{a_n a_{n+t}} - \expect{a_n}^2}{\expect{a_n^2} - \expect{a_n}^2} . \label{equation:autocorrelation-definition}
\end{eqnarray}
The basic concept is that $\tau_\mathrm{eq}$ corresponds to the single-exponential decay time for the autocorrelation process that generates samples, so the statistical inefficiency $g$ measures the approximate temporal separation between two effectively uncorrelated samples (where two exponential relaxation times are presumed to be sufficient).

Robust estimation of $C_t$ for $t \sim T$ is difficult due to growth in statistical error, so common estimators of $g$ make use of several additional properties of $C_t$ to provide useful estimates (see \emph{Practical Computation of Statistical Inefficiencies} in~\cite{chodera2016simple} for a detailed discussion).
\todo[inline]{JDC: Describe recommended statistical inefficiency computation algorithm.}
We recommend using the robust statistical inefficinecy computation routines available within the MBAR~\cite{kylebeauchamp2019choderalab} and alchemlyb~\cite{daviddotson2019alchemistry} packages.

\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can be fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_{\mathrm{eff}} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \mathrm{range}(1,\ldots ,N) \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.

If independent simulations are used, the alchemical state $\lambda$ may have a significant impact on the correlation time, and these simulations should be subsampled independently using a separate estimate of the statistical inefficinecy $g$ for each alchemical state.
If coupled simulations are used (such as a Hamiltonian replica exchange simulation), the replicas should undergo equivalent random walks in alchemical space, and the replicas can be can be subsampled with the same $g$ to generate an equal number of uncorrelated samples at each alchemical state.
Conveniently, the approach described above for automated equilibration detection produces an appropriate estimate of $g$ over the production region for automating this process.

\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_{\mathrm{eff}} \ge \approx 50$ if the BAR or MBAR estimators (discussed below) are used; the number may need to be much higher with alternate estimators.

@@

Free energy differences between two different states differing in the energy function are directly related to the
ratio of probabilities of those states.
As can be noted, the partition functions in Eq.~\ref{equation:dimensionless-free-energy-difference} are simply the total accumulated probabilites for all possible configurations of the system. Virtually all of the ways to estimate this free energy are based in converting this ratio of integrals to something that can be measured in one (or several) simulations.

@@

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought this final term is sometimes used to encompass all ways of calculating free energy differences.

The (reduced) free energy difference $\Delta f_{01}$ between an initial state 0 and a final state 1 defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - u_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta u(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
and the average is over all samples from the simulation performed with $u_0$. In the case of NVT (canonical) sampling and assuming the masses do not change, then $u$ is just $U/k_BT$, and $f$ is $A/k_BT$, but it can be generalized to other ensembles with the proper definition of $f$ and $u$.
Described in words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be if we switched to potential energy function $U_1(\vec{q})$, and average the exponential of the -1 times the energy difference to get -1 times the exponential of the free energy difference.   This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies of stored snapshots.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change under the same equilibrium conditions at either end state~\cite{jarzynski1997nonequilibrium,jarzynski1998equilibrium,crooks2000pathensemble}, though we do not go into all of the details of nonequilibrum transformations here, and refer the reader to more advanced treatments~\cite{maragakis2008bayesian,oberhofer2005biased,procacci2015unbiased,shirts2003equilibriuma,ytreberg2004singleensemble}

Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume, which is true for standard molecular models), it has some very important numerical issues that mean that it often performs badly for standard free energy calculations, even for small molecules.~\cite{shirts2005comparison,lu2003appropriate}  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{xx}. Essentially, the method is dominated by contributions of rare snapshots\cite{jarzynski2006rare, wu2005phasespaceb, wu2005phasespacec}. There are a number of applications where it this expression is nevertheless useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}

@@

If we have the differences in the potential energy sampled from the distribution defined by $U_0$ to the state defined by $U_1$, and we also have the differences in potential energies from the distribution sampled by $U_1$ to the state defined by $U_0$, we can obtain a significantly improved estimate of the
free energy difference compared to that obtained by EXP.  
This estimate was first derived by Bennett and is hence generally called the Bennett Acceptance Ratio (BAR).  It is solved by finding the reduced free energy $f_{ij}$ that satisfied the following implicit equation:
\begin{eqnarray}
 \sum_{i=1}^{n_i} \frac{1}{1 + \exp[\ln(\frac{n_i}{n_j}) + u_{ij}(\vec{q}) - f_{ij})
 ]} \nonumber \\
 =\sum_{i=1}^{n_j} \frac{1}{1 + \exp[\ln(\frac{n_i}{n_j}) - U_{ij}(\vec{q}) + f_ij)]}
\end{eqnarray}
where $n_i$ and $n_j$ are the number of samples from each state. We
will refer to this method as BAR. More recent derivations show that this formula is the maximum likelihood estimate of the free energy difference
given sets of samples from the two states~\cite{shirts.bennett}.

Many studies have demonstrated both the
theoretical and practical superiority of BAR over EXP in molecular
simulations~\cite{shirts2005comparison,lu2003appropriate}, and BAR converges to EXP in the limit that all samples are from a
single state~\cite{bennett1976efficient,shirts.bennett}. BAR also requires significantly less
overlap between the configurational space of each state to converge than EXP, though some overlap must still exist.

The Bennett acceptance ratio is only defined between two states.  Usually, the endpoints of interest in a free energy calculation are sufficiently different that we will need a chain of states that gradually change the potential energy function from $U_0$to $U_1$, as discussed in Section~\ref{}. One can simply carry out BAR between each pair of states $\Delta G_{1 \rightarrow N} = \Delta {G_{1\rightarrow 2}} + \Delta {G_{2\rightarrow 3}} +  \ldots + \Delta G_{N-1\rightarrow N}$.

There is one important thing to note about the uncertainty estimates when summing multiple free energies together to calculate an overall free energy estimate.  Although BAR itself gives a free energy estimate that is asymptotically correct in and is much less biased than the uncertainty estimate for EXP, the uncertainties in $\Delta {G_{i-1\rightarrow i}}$ and $\Delta {G_{i\rightarrow i+1}}$ are not uncorrelated, because they both involve the energies $U_i(\vec{q})$. The variances of each of the free energies will \textit{not} propagate as variances usually do (in quadrature) to the variance of the overall free energy. Instead, some other method for propagating the uncertainty, such as bootstrapping~\cite{Grossfield:L:2018} must be used.

@@

By taking the derivative of the free energy with respect to the
variable $\lambda$, we find that:
\begin{equation}
\frac{df}{d\lambda} = \frac{d}{d\lambda} \ln \int \exp{-u(\lambda,\vec{q})} d\vec{q} = \expect{\frac{du(\lambda,\vec{q})}{d\lambda}\
}_{\lambda} 
\end{equation}

And than we can then numerically integrate $df/d\lambda$ over an alchemical transformation, using a range of different well-established techniques, to obtain:
\begin{equation}
\Delta f    = \int_{0}^{1} \expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{\lambda}  d\lambda    
\end{equation}
This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\lambda}}$ requires
fewer uncorrelated samples to reach a given level of relative error
than averaging $e^{-u(\vec{q})}$, as the range is usually
narrower with a more Gaussian distribution. Rather than being limited by overlap, as in the case of BAR and MBAR, we are instead limited by the bias in the numerical quadrature, which must be minimized sufficiently to be beneath the level of statistical noise.

Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme.  All types of
numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{k} \] where the weights
$w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration
schemes~\cite{resat1993studies,jorge2010effect,shyu2009reducing}. However, many other integration choices require specific choices of $\lambda$
to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligble benefits~\cite{paliwal2011benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\lambda$ are simulated.  Because fitting to higher
order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate
with some types of transformations, expertise and experience is
required to perform such numerical integration modifications.  In practice, adding 2-3 more intermediate states is typically sufficient to match the performance of these more complicated numerical quadrature schemes.

One drawback of TI is that it requires derivatives with respect to $\lambda$ to be calculated directly in the code. Unfortunately, many problems of interest
require using pathways (such as the soft-core pathways, for removing repulsive interactions) that are not linear, as we discuss, making this more complex.  Still, if the code of interest does compute $\frac{du}{d\lambda}$, then TI is perhaps the simplest method to use, as it involves a very little post-processing,

@@

One can generalize Bennett's logic from two states to multiple states to obtain a free energy estimator that uses energy differences between all intermediate states.  MBAR gives a system of implicit equations for the free energies $f_i$:
\begin{equation}
f_i = - \ln \sum_{n=1}^{N} \frac{\exp(-u_i(\vec{q}_n))}{\sum_{k=1}^K N_k \exp(f_k-u_k(\vec{q}_n))}
\end{equation}
where there are $N_k$ samples from each of $K$ states, with $\sum_k N_k=N$ the total number of samples. Thus, we need to evaluate the energy function $U_i$ for all samples obtained at all states in the transformation. The equations can be solved by a number of different standard routines.  We note that there are only $K-1$ independent equations, so only K-1 of the free energies are independent variables, and one of the $f_i$ must be specified (usually, without loss of generality, setting it to zero).

>>>>

In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of alchemical calculations~(\ref{sec:data_analysis}). 

@@

\label{subsec:expectation}

@@

Based on the input ligand series a perturbation map or network can be planned. Recent heuristics have shown the more connected the perturbation network the better, however, there is a way to optimise network structure while minimising the number of perturbations that need to be computed minimising the resulting computational cost~\cite{yang2020optimal}. Sometimes the introduction of intermediates that are not part of the original congeneric series are essential to avoid ring breaking, or dealing with perturbations that would otherwise result in large numbers of atoms to be grown or disappearing. Some commercial tools have underlying good heuristic but may fail with complicated input, needing user validation in particular when dealing with chiral compounds. 

@@

Methods in experimental design have been applied to the construction of the perturbation maps.  Yang et al.~\cite{yang2020optimal} optimized the perturbation map by selecting a fixed number of calculations from the pairwise perturbations so that the resulting set of calculations minimize the total variance.  Xu~\cite{xu2019diffnet} optimized the perturbation map by allocating different amounts of simulation time to different pairwise perturbations so as to minimize the total variance, given the total simulation time of all the perturbation calculations.  Both approaches lead to substantial reduction in the statistical error of the estimated free energies.  

@@

\subsubsection{The importance of the alchemical pathway
\label{sec:important_path}}

@@

Free energy information should generally be saved more frequently than coordinate data, approximately at the rate that uncorrelated samples are produced.  
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected frequently. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. One may safely assume that the correlation time is greater than 100-200 fs even for relatively simple systems such as small molecules in solvent, so saving no more frequently than every 50-100 steps is recommended. 
How decorrelation impacts calculations, and how to compute it is discussed in other sections.  [section ref]
%
\section{Data analysis}
\label{sec:data_analysis}
Once data has been collected from alchemical intermediates, it must be analyzed to produce an estimate of the free energy change (and its associated statistical uncertainty) for each leg of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to provide information on poor convergence.
%
\subsection{Detecting the boundary between equilibrated and production regions}
\label{sec:automatic-equilibration-detection}
Much of the infrastructure for analyzing alchemical free energy calculations relies on the concept of asymptotically unbiased estimators, which produce unbiased estimates of the free energy when fed very long simulations~\cite{shirts2005comparison}.
In reality, free energy calculations are often initiated from highly atypical initial conditions (such as a protein-ligand geometry obtained from docking and subjected to a heuristic solvent placement scheme), and simulations are of a finite length dictated by available computational resources and computing demands.
As a result, these estimators can produce significantly biased estimates if fed the entirety of simulation data generated without further processing~\cite{chodera2016simple}.
%
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/fig11_equib_detection/Figure.pdf}
    \caption{{\bf Automatic partitioning into equilibration and production regions.}
    \textbf{A} The average (black line) standard deviation (shaded region) of the reduced potential $u^*$ over many independent replicate simulations started from the same initial conditions show a significant initial transient before relaxing to the true average potential energy\textbf{B} A cumulative average (red) of the entire simulation data demonstrates simulation bias not seen when initial simulation data is omitted (blue). Using an automated approach to detect equilibration of the boundary $t_0$ using statistical inefficiency $g$ (\textbf{C})  for an effective simulation interval (\textbf{D}) \textbf{E} The optimal equilibration boundary $t_0$ is selected to maximize the number of uncorrelated samples.
    \emph{Figure adapted from~\cite{chodera2016simple}.}
    }
    \label{fig:automatic-equilibration-detection}
\end{figure}

@@

\paragraph{Cautions in automating equilibration detection}
For simulations that are simply not long enough to contain a large number of samples from true equilibrium either because they are very short or contain slow processes, this procedure cannot completely remove the bias.
In such cases, this approach simply selects the final portion of the the simulation, which may be contained in a single substate of conformational space, which may itself lead to biased estimates. 
This situation can be detected if the equilibration boundary $t_0$ is a significant fraction of the total simulation length $T$, with a good rule of thumb being that $T \gtrsim 20 t_0$.
If this is not possible, advanced analysis techniques that assume only local equilibrium (rather than global equilibrium) such as the TRAM estimators~\cite{mey2014xtram,wu2016multiensemble,nuske2017markov} may be more appropriate.
%
\subsection{Decorrelating samples for analysis}
\label{sec:decorrelating-samples}
\paragraph{Computing the statistical inefficiency}
Most estimators require an uncorrelated set of samples from the equilibrium distribution to produce (relatively) unbiased estimates of the free energy difference and its statistical uncertainty.
To do this, the production region of the simulation is generally \emph{subsampled} with an interval approximately equal to or greater than the \emph{statistical inefficiency} $g \ge 1$ to produce a set of uncorrelated samples that can be fed to the estimator machinery~\cite{chodera2016simple},
\begin{eqnarray}
g &\equiv& 1 + 2 \tau_\mathrm{eq} \label{eq:statistical-inefficiency-definition}
\end{eqnarray}
where $\tau_\mathrm{eq}$ is the integrated autocorrelation time, formally defined as
\begin{eqnarray}
\tau_\mathrm{eq} &\equiv& \sum_{t=1}^{T-1} \left(1 - \frac{t}{T}\right) C_t \label{eq:integrated-autocorrelation-time-definition} , 
\end{eqnarray}
with the discrete-time normalized fluctuation autocorrelation function $C_t$ defined as
\begin{eqnarray}
C_t &\equiv& \frac{\expect{a_n a_{n+t}} - \expect{a_n}^2}{\expect{a_n^2} - \expect{a_n}^2} . \label{equation:autocorrelation-definition}
\end{eqnarray}
The basic concept is that $\tau_\mathrm{eq}$ corresponds to the single-exponential decay time for the autocorrelation process that generates samples, so the statistical inefficiency $g$ measures the approximate temporal separation between two effectively uncorrelated samples (where two exponential relaxation times are presumed to be sufficient).

Robust estimation of $C_t$ for $t \sim T$ is difficult due to growth in statistical error, so common estimators of $g$ make use of several additional properties of $C_t$ to provide useful estimates (see \emph{Practical Computation of Statistical Inefficiencies} in~\cite{chodera2016simple} for a detailed discussion).
\todo[inline]{JDC: Describe recommended statistical inefficiency computation algorithm.}
We recommend using the robust statistical inefficinecy computation routines available within the MBAR~\cite{kylebeauchamp2019choderalab} and alchemlyb~\cite{daviddotson2019alchemistry} packages.
\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can be fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_{\mathrm{eff}} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \mathrm{range}(1,\ldots ,N) \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.
If independent simulations are used, the alchemical state $\lambda$ may have a significant impact on the correlation time, and these simulations should be subsampled independently using a separate estimate of the statistical inefficinecy $g$ for each alchemical state.
If coupled simulations are used (such as a Hamiltonian replica exchange simulation), the replicas should undergo equivalent random walks in alchemical space, and the replicas can be can be subsampled with the same $g$ to generate an equal number of uncorrelated samples at each alchemical state.
Conveniently, the approach described above for automated equilibration detection produces an appropriate estimate of $g$ over the production region for automating this process.
\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_{\mathrm{eff}} \ge \approx 50$ if the BAR or MBAR estimators (discussed below in sec~\ref{subsec:estimators}) are used; the number may need to be much higher with alternate estimators.

@@

\label{subsec:estimators}
Free energy differences between two different states differing in the energy function are directly related to the
ratio of probabilities of those states.
As can be noted, the partition functions in Eq.~\ref{equation:dimensionless-free-energy-difference} are simply the total accumulated probabilites for all possible configurations of the system. Virtually all of the ways to estimate this free energy are based in converting this ratio of integrals to something that can be measured in one (or several) simulations.  

@@

The simplest method for calculating free energy differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought this final term is sometimes used to encompass all ways of calculating free energy differences.
The (reduced) free energy difference $\Delta f_{01}$ between an initial state 0 and a final state 1 defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - u_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta u(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
and the average is over all samples from the simulation performed with $u_0$. In the case of NVT (canonical) sampling and assuming the masses do not change, then $u$ is just $U/k_BT$, and $f$ is $F/k_BT$, but it can be generalized to other ensembles with the proper definition of $f$ and $u$.
Described in words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be if we switched to potential energy function $U_1(\vec{q})$, and average the exponential of the -1 times the energy difference to get -1 times the exponential of the free energy difference. The original distributions (P($U_0$) as generated at $\lambda=0$ and P($U_1$) would look like the ones seen in figure~\ref{fig:fig_sampling_scheme} (A)-(C)on the left hand side. Reevaluating  requires almost no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies of stored snapshots. The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change under the same equilibrium conditions at either end state~\cite{jarzynski1997nonequilibrium,jarzynski1998equilibrium,crooks2000pathensemble}, though we do not go into all of the details of non-equilibrium transformations here, and refer the reader to more advanced treatments~\cite{maragakis2008bayesian,oberhofer2005biased,procacci2015unbiased,shirts2003equilibriuma,ytreberg2004singleensemble}
Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume, which is true for standard molecular models), it has some very important numerical issues that mean that it often performs badly for standard free energy calculations, even for small molecules~\cite{shirts2005comparison,lu2003appropriate}.  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{lelievre2010free}. Essentially, the method is dominated by contributions of rare snapshots\cite{jarzynski2006rare, wu2005phasespaceb, wu2005phasespacec}. There are a number of applications where it this expression is nevertheless useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}

@@

If we have the differences in the potential energy sampled from the distribution defined by $U_0$ to the state defined by $U_1$, and we also have the differences in potential energies from the distribution sampled by $U_1$ to the state defined by $U_0$, we can obtain a significantly improved estimate of the
free energy difference compared to that obtained by EXP. 
This estimate was first derived by Bennett and is hence generally called the Bennett Acceptance Ratio (BAR).  It is solved by finding the reduced free energy $f_{ij}$ that satisfied the following implicit equation:
\begin{eqnarray}
 \sum_{i=1}^{n_i} \frac{1}{1 + \exp[\ln(\frac{n_i}{n_j}) + u_{ij}(\vec{q}) - f_{ij})
 ]} \nonumber \\
 =\sum_{i=1}^{n_j} \frac{1}{1 + \exp[\ln(\frac{n_i}{n_j}) - U_{ij}(\vec{q}) + f_ij)]},
\end{eqnarray}
where $n_i$ and $n_j$ are the number of samples from each state. We
will refer to this method as BAR. More recent derivations show that this formula is the maximum likelihood estimate of the free energy difference given sets of samples from the two states~\cite{bennett1976efficient,shirts2003equilibriuma }. 
Many studies have demonstrated both the theoretical and practical superiority of BAR over EXP in molecular
simulations~\cite{shirts2005comparison,lu2003appropriate}, and BAR converges to EXP in the limit that all samples are from a single state~\cite{bennett1976efficient,bennett1976efficient,shirts2003equilibriuma}. BAR also requires significantly less overlap between the configurational space of each state to converge than EXP, though some overlap must still exist.
The Bennett acceptance ratio is only defined between two states. Usually, the endpoints of interest in a free energy calculation are sufficiently different that we will need a chain of states that gradually change the potential energy function from $U_0$to $U_1$, as discussed in Section~\ref{sec:important_path}. You can simply carry out BAR between each pair of states $\Delta G_{1 \rightarrow N} = \Delta {G_{1\rightarrow 2}} + \Delta {G_{2\rightarrow 3}} +  \ldots + \Delta G_{N-1\rightarrow N}$.
There is one important thing to note about the uncertainty estimates when summing multiple free energies together to calculate an overall free energy estimate. Although BAR itself gives a free energy estimate that is asymptotically correct in and is much less biased than the uncertainty estimate for EXP, the uncertainties in $\Delta {G_{i-1\rightarrow i}}$ and $\Delta {G_{i\rightarrow i+1}}$ are not uncorrelated, because they both involve the energies $U_i(\vec{q})$. The variances of each of the free energies will \textit{not} propagate as variances usually do (in quadrature) to the variance of the overall free energy. Instead, some other method for propagating the uncertainty, such as bootstrapping~\cite{grossfield2018best} must be used.

@@

By taking the derivative of the free energy with respect to the
variable $\lambda$, we find that:
\frac{df}{d\lambda} = \frac{d}{d\lambda} \ln \int \exp{-u(\lambda,\vec{q})} d\vec{q} = \expect{\frac{du(\lambda,\vec{q})}{d\lambda}\
}_{\lambda} 
And than we can then numerically integrate $df/d\lambda$ over an alchemical transformation, using a range of different well-established techniques, to obtain:
\Delta f    = \int_{0}^{1} \expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{\lambda}  d\lambda    
This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\lambda}}$ requires fewer uncorrelated samples to reach a given level of relative error
than averaging $e^{-u(\vec{q})}$, as the range is usually narrower with a more Gaussian distribution. Rather than being limited by overlap, as in the case of BAR and MBAR (see below), we are instead limited by the bias in the numerical quadrature, which must be minimized sufficiently to be beneath the level of statistical noise.
Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme.  All types of numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{k} \] where the weights $w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration schemes~\cite{resat1993studies,jorge2010effect,shyu2009reducing}. However, many other integration choices require specific choices of $\lambda$ to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligible benefits~\cite{paliwal2011benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\lambda$ are simulated.  Because fitting to higher order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate with some types of transformations, expertise and experience is required to perform such numerical integration modifications. In practice, adding 2-3 more intermediate states is typically sufficient to match the performance of these more complicated numerical quadrature schemes.
One drawback of TI is that it requires derivatives with respect to $\lambda$ to be calculated directly in the code. Unfortunately, many problems of interest require using pathways (such as the soft-core pathways, for removing repulsive interactions) that are not linear, as we discuss, making this more complex. Still, if the code of interest does compute $\frac{du}{d\lambda}$, then TI is perhaps the simplest method to use, as it involves a very little post-processing.

@@

One can generalize Bennett's logic from two states to multiple states to obtain a free energy estimator that uses energy differences between all intermediate states. MBAR gives a system of implicit equations for the free energies $f_i$:
f_i = - \ln \sum_{n=1}^{N} \frac{\exp(-u_i(\vec{q}_n))}{\sum_{k=1}^K N_k \exp(f_k-u_k(\vec{q}_n))}
where there are $N_k$ samples from each of $K$ states, with $\sum_k N_k=N$ the total number of samples. Thus, we need to evaluate the energy function $U_i$ for all samples obtained at all states in the transformation. The equations can be solved by a number of different standard routines. We note that there are only $K-1$ independent equations, so only K-1 of the free energies are independent variables, and one of the $f_i$ must be specified (usually, without loss of generality, setting it to zero).
